Catching fire: Undefined behaviour for equational reasoning of lazy languages with strictness annotations

### 2021-01-21 paper/paper.tex
add explanation of the simple example

@@

%\begin{figure}[htb!]
%\begin{minted}{cpp}
{\footnotesize
\begin{lstlisting}[caption=Step 1: outlining everything after the initial \lzforce into \texttt{f2}]
SimpleInt f2(SimpleInt icons) {
    int ihash = casedefault(icons);
    if (ihash <= 0) {
        return SimpleInt(42);
    } else {
        int prev = ihash - 1;
        SimpleInt siprev = SimpleInt(prev);
        Thunk<SimpleInt> siprev_t = thunkify(siprev);
        SimpleInt f_prev_v = apStrict(f, siprev_t);
        return f_prev_v;
    }
}

SimpleInt f(Thunk<SimpleInt> i) {
    SimpleInt icons = force(i);
    SimpleInt ret = apStrict(f2, icons);
    return ret;
}
\end{lstlisting}
}
%\caption{Step 1: outlining everything after the initial \lzforce into \texttt{f2}}
%\end{figure}

%\begin{figure}[htb!]
%\begin{minted}{cpp}
\begin{lstlisting}[language=c++, caption=Step 2: Removing the strict application]
SimpleInt f2(SimpleInt icons) {
    int ihash = casedefault(icons);
    if (ihash <= 0) {
        return SimpleInt(42);
    } else {
        int prev = ihash - 1;
        SimpleInt siprev = SimpleInt(prev);
        // Thunk<SimpleInt> siprev_t = thunkify(siprev);
        // SimpleInt f_prev_v = apStrict(f, siprev_t);
        Thunk<SimpleInt> f_prev_v = apStrict(f2, siprev);
        return f_prev_v;
    }
}

SimpleInt f(Thunk<SimpleInt> i) {
    SimpleInt icons = force(i);
    SimpleInt ret = apStrict(f2, icons);
    return ret;
}
\end{lstlisting}
%\end{minted}
%\caption{Step 2: Removing the strict application}
%\end{figure}

%\begin{figure}[htb!]
{\footnotesize
\begin{lstlisting}[language=c++, caption=Step 3: outline everything after casedefault into f3]
SimpleInt f3(int icons) {
    if (ihash <= 0) {
        return SimpleInt(42);
    } else {
        int prev = ihash - 1;
        SimpleInt siprev = SimpleInt(prev);
        Thunk<SimpleInt> f_prev_v = apStrict(f2, siprev);
        return f_prev_v;
    }
}

SimpleInt f2(SimpleInt icons) {
    int ihash = casedefault(icons);
    SimpleInt ret = apStrict(f3, ihash);
    return ret;
}

SimpleInt f(Thunk<SimpleInt> i) {
    SimpleInt icons = force(i);
    SimpleInt ret = apStrict(f2, icons);
    return ret;
}
\end{lstlisting}
}
%\caption{ Step 3: outline everything after casedefault(icons) into f3}
%\end{figure}

%\begin{figure}[htb!]
{\footnotesize
\begin{lstlisting}[language=c++, caption=replace call \texttt{f2(SimpleInt(prev))} to \texttt{f3(prev)}]
SimpleInt f3(int icons) {
    if (ihash <= 0) {
        return SimpleInt(42);
    } else {
        int prev = ihash - 1;
        Thunk<SimpleInt> f_prev_v = apStrict(f3, prev);
        return f_prev_v;
    }
}
SimpleInt f2(SimpleInt icons) {
    int ihash = casedefault(icons);
    SimpleInt ret = apStrict(f3, ihash);
    return ret;
}
SimpleInt f(Thunk<SimpleInt> i) {
    SimpleInt icons = force(i);
    SimpleInt ret = apStrict(f2, icons);
    return ret;
}
\end{lstlisting}
}
%\caption{Step 4: replace call f2(SimpleInt(prev)) to f3(prev)}
%\end{figure}

>>>>

\lstset{
language=C++,
basicstyle=\small\ttfamily,
escapeinside={<@}{@>},
columns=fullflexible,
keywords=[2]{old},
keywords=[3]{new},
keywordstyle={\color{blue!80!black}},
keywordstyle=[2]{\color{red!80!black}},
keywordstyle=[3]{\color{green!50!black}},
}

\colorlet{diffold}{red!80!black}
\colorlet{diffnew}{green!50!black}

@@

%\begin{figure}[htb!]
%\begin{minted}{cpp}
{\footnotesize
\begin{lstlisting}[caption=Step 1: outlining everything after the initial \lzforce into \texttt{f2}]
SimpleInt f2(SimpleInt icons) {
    int ihash = casedefault(icons);
    if (ihash <= 0) {
        return SimpleInt(42);
    } else {
        int prev = ihash - 1;
        SimpleInt siprev = SimpleInt(prev);
        Thunk<SimpleInt> siprev_t = thunkify(siprev);
        SimpleInt f_prev_v = f(siprev_t);
        return f_prev_v;
    }
}
SimpleInt f(Thunk<SimpleInt> i) {
    SimpleInt icons = force(i);
    SimpleInt ret = f2(icons);
    return ret;
}
\end{lstlisting}
}
%\caption{Step 1: outlining everything after the initial \lzforce into \texttt{f2}}
%\end{figure}

%\begin{figure}[htb!]
%\begin{minted}{cpp}
\begin{lstlisting}[language=c++, caption=Step 2: Removing the un-necessary lazy recursive call]
SimpleInt f2(SimpleInt icons) {
    int ihash = casedefault(icons);
    if (ihash <= 0) {
        return SimpleInt(42);
    } else {
        int prev = ihash - 1;
        SimpleInt siprev = SimpleInt(prev);
        <@\textcolor{diffold}{Thunk<SimpleInt> siprev\_t = thunkify(siprev);}@>
        <@\textcolor{diffold}{SimpleInt f\_prev\_v = f(siprev\_t);} @>
       <@ \textcolor{diffnew}{Thunk<SimpleInt> f\_prev\_v = f2(siprev);} @>
        return f_prev_v;
    }
}

SimpleInt f(Thunk<SimpleInt> i) {
    SimpleInt icons = force(i);
    SimpleInt ret = f2(icons);
    return ret;
}
\end{lstlisting}
%\end{minted}
%\caption{Step 2: Removing the strict application}
%\end{figure}

%\begin{figure}[htb!]
{\footnotesize
\begin{lstlisting}[language=c++, caption=Step 3: outline everything after casedefault into f3]
SimpleInt f3(int icons) {
    if (ihash <= 0) {
        return SimpleInt(42);
    } else {
        int prev = ihash - 1;
        SimpleInt siprev = SimpleInt(prev);
        Thunk<SimpleInt> f_prev_v = f2(siprev);
        return f_prev_v;
    }
}

SimpleInt f2(SimpleInt icons) {
    int ihash = casedefault(icons);
    <@ \textcolor{diffnew}{SimpleInt ret = f3(ihash);} @>

    return ret;
}
...
\end{lstlisting}
}
%\caption{ Step 3: outline everything after casedefault(icons) into f3}
%\end{figure}

%\begin{figure}[htb!]
{\footnotesize
\begin{lstlisting}[language=c++, caption=replace call \texttt{f2(SimpleInt(prev))} to \texttt{f3(prev)}]
SimpleInt f3(int icons) {
    if (ihash <= 0) {
        return SimpleInt(42);
    } else {
        int prev = ihash - 1;
        <@ \textcolor{diffold}{SimpleInt siprev = SimpleInt(prev);} @>
        <@ \textcolor{diffold}{Thunk<SimpleInt> f\_prev\_v = f2(siprev);} @>
        <@ \textcolor{diffnew}{Thunk<SimpleInt> f\_prev\_v = f3(prev);} @>
        return f_prev_v;
    }
}
...
\end{lstlisting}
}

%\caption{Step 4: replace call f2(SimpleInt(prev)) to f3(prev)}
%\end{figure}

{\footnotesize
\begin{lstlisting}[language=c++, caption=taking stock]
SimpleInt f3(int icons) {
    if (ihash <= 0) {
        return SimpleInt(42);
    } else {
        int prev = ihash - 1;
        <@ \textcolor{diffnew}{Thunk<SimpleInt> f\_prev\_v = f3(prev);} @>
        return f_prev_v;
    }
}

SimpleInt f2(SimpleInt icons) {
    int ihash = casedefault(icons);
    <@ \textcolor{diffnew}{SimpleInt ret = f3(ihash);} @>
    return ret;
}
SimpleInt f(Thunk<SimpleInt> i) {
    SimpleInt icons = force(i);
    SimpleInt ret = f2(icons);
    return ret;
}
\end{lstlisting}
}

At this point, we have a clean program that has been worker/wrappered. We can
see that the call chain looks as follows:

$$
f \xrightarrow{\texttt{force}} f_2 \xrightarrow{\texttt{unwrap}} f_3 \to f_3 \dots \xrightarrow{\texttt{wrap}} f_2
\xrightarrow{\texttt{thunk}} f
$$
